---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r, echo=FALSE}
# loading packages
library(plyr)
library(dplyr)
```


## Loading and preprocessing the data
First, we want to load and preprocess the data. The data in included in this repo zipped up. We unzip activity.zip and load it into data.

```{r, echo=TRUE}
# unzip data
if (!file.exists("activity.csv")) {
    unzip("activity.zip", "activity.csv")
}
# load data
data <- read.csv("activity.csv")
summary(data)
```

Then we preprocess:
```{r, echo=TRUE}
# preprocess
data$date <- as.Date(data$date, format = "%Y-%m-%d")

# data to answer total number of steps per day question
d.steps <- data %>%
           filter(!is.na(steps)) %>% 
           group_by(date) %>% 
           summarise(total_steps = sum(steps, na.rm = TRUE))

# data to answer daily pattern question
d.interval <- data %>% 
              filter(!is.na(steps)) %>%                          
              group_by(interval) %>%         
              summarise(steps_mean = mean(steps))

```

## What is mean total number of steps taken per day?

Calculate total number of steps taken per day:
```{r, echo=TRUE}
print(d.steps)
```

Then a histogram:

```{r, echo=TRUE}
hist(d.steps$total_steps, main="Histogram of total steps per day", xlab = "Total steps per day")
```

Finally here are mean and median of total number of steps per day:

```{r, echo=TRUE}
steps_mean <- mean(d.steps$total_steps)
steps_mean

steps_median <- median(d.steps$total_steps)
steps_median
```

## What is the average daily activity pattern?

```{r, echo=TRUE}

plot(d.interval$interval, d.interval$steps_mean, type = "l", xlab = "interval", ylab = "steps", main = "Time series of steps against interval")
```

Interval with most number of steps:
```{r, echo=TRUE}
d.interval[which.max(d.interval$steps_mean),]
```

## Imputing missing values
First we count the total number of NAs in dataset:
```{r, echo=TRUE}
colSums(is.na(data))
```


We create a new dataset d.fill_missing and fill missing values with average number of steps in 5-min interval:

```{r, echo=TRUE}
d.fill_missing  <- data
missing <- is.na(d.fill_missing$steps)
mean_interval <- tapply(d.fill_missing$steps, d.fill_missing$interval, mean, na.rm = TRUE)

d.fill_missing$steps[missing] <- mean_interval[as.character(d.fill_missing$interval[missing])]
```

As expected we find no NAs in this new data set:

```{r, echo=TRUE}
colSums(is.na(d.fill_missing))

```


## Are there differences in activity patterns between weekdays and weekends?
